Learning Libraries of Subroutines for Neurally--Guided Bayesian Program Learning

### 2018-05-18 nips_2018.tex
text editing figure

  \hspace{0.5cm}\begin{tabular}{l}
    $f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\code{(map ($\lambda$ (x)}\\\hspace{1cm}\code{ (if (= x a) b x)) s)}\\
      \hspace{\helpSize}($f_0$: \emph{Performs character substitution)}\\
      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a)}\\\hspace{1.1cm}\code{ (cdr (if (= c x) s a))))}\\
        \hspace{\helpSize}($f_1$: \emph{Drop characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
      $f_2($\code{s}$) \,=\, $\code{(unfold s is-nil car }\\\hspace{1cm}\code{($\lambda$ (z) (}$f_1$\code{ z \codechar{ })))}\\
        \hspace{\helpSize}($f_2$: \emph{Abbreviates a sequence of words})\\
  \end{tabular}&

@@

Because our enumerative search procedure has no means of generating
string constants, we have the enumerator propose programs with string-valued holes --
e.g., in Fig.~\ref{firstPageFigure},
we enumerate $\text{\code{(}}f_0\code{ string (}f_2\text{\code{ s ' '))}}$ --
and define $\probability[x|p]$ by marginalizing out the values of the
strings via dynamic programming.  In Section~\ref{regressionSection},
we will use a similar trick to synthesize programs containing real
numbers, but using gradient descent instead of dynamic programming.

We automatically generated 109 text editing tasks with 4 input/output examples each
%(Fig.~\ref{exampleTextProblem}).
At first, \system cannot find any correct programs for most of the tasks.
After three iterations, it assembles a DSL containing 12 new functions (center of Fig.~\ref{initialExampleDSL}) that lets it rapidly explore the space of programs and find solutions to
all of the tasks.

How well does the  learned DSL generalized to real text-editing scenarios?
We tested, but did not train, our system on the 108 text editing problems from the SyGuS~\cite{alur2016sygus} program synthesis competition. Before any learning,
\system solves 3.7\% of the problems with an average search time of 235 seconds.
After learning,
it solves 74.1\%, and does so much faster,
solving them in an average of 38 seconds.
As of the 2017 SyGuS competition,
the best-performing algorithm solves 79.6\% of the problems.
But, SyGuS comes with a
different hand-engineered DSL \emph{for each text editing problem}.
Here  we learned a single DSL
that applied generically to
all of the tasks,
and perform comparably to the best
prior work.

@@

  \\\midrule\multicolumn{7}{c}{\emph{Text Editing}}\\\midrule
  \% solved&\textbf{74\%} &43\% &30\%&33\%&0\%&4\%\\
  Solve time&  38s&65s&38s&80s&--&235s

>>>>

  \hspace{0.5cm}\begin{tabular}{l}
    $f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\code{(map ($\lambda$ (x)}\\\hspace{1cm}\code{ (if (= x a) b x)) s)}\\
      \hspace{\helpSize}($f_0$: \emph{Performs character substitution)}\\
      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(foldr s s ($\lambda$ (x a)}\\\hspace{1.1cm}\code{ (cdr (if (= c x) s a))))}\\
        \hspace{\helpSize}($f_1$: \emph{Drop characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
      $f_2($\code{s}$) \,=\, $\code{(unfold s is-nil car }\\\hspace{1cm}\code{($\lambda$ (z) (}$f_1$\code{ z \codechar{ })))}\\
        \hspace{\helpSize}($f_2$: \emph{Abbreviates a sequence of words})\\
        $f_3($\code{a}$,$\code{b}$) \,=\, $\code{(foldr a b cons)}\\
      \hspace{\helpSize}($f_3$: \emph{Concatenate strings }\code{a}\emph{ and }\code{b})
  \end{tabular}&

@@

Because our enumerative search procedure cannot generate string
constants, we instead enumerate programs with string-valued
parameters.  For example, to learn a program that prepends ``Dr.'', we
enumerate $\text{\code{(}}f_3\code{ string s)}$ -- where $f_3$ is the
learned appending primitive (Fig.~\ref{initialExampleDSL}) -- and then
define $\probability[x|p]$ by approximately marginalizing out the
string parameters via a simple dynamic program.
In Sec.~\ref{regressionSection}, we will use a similar trick to
synthesize programs containing real numbers, but using gradient
descent instead of dynamic programming.

We trained our system on a corpus of 109 automatically generated text editing tasks, with 4 input/output examples each.
After three iterations, it assembles a DSL containing a dozen new functions (center of Fig.~\ref{initialExampleDSL}) that let it solve 
all of the training tasks.

How well does the  learned DSL generalized to real text-editing scenarios?
We tested, but did not train, on the 108 text editing problems from the SyGuS~\cite{alur2016sygus} program synthesis competition. Before any learning,
\system solves 3.7\% of the problems with an average search time of 235 seconds.
After learning,
it solves 74.1\%, and does so much faster,
solving them in an average of 29 seconds.
As of the 2017 SyGuS competition,
the best-performing algorithm solves 79.6\% of the problems.
But, SyGuS comes with a
different hand-engineered DSL \emph{for each text editing problem}.\footnote{SyGuS text editing problems also prespecify the set of allowed string constants \emph{for each task}. Our system does not use this assistance.}
Here  we learned a single DSL
that applied generically to
all of the tasks,
and perform comparably to the best
prior work.

@@

  \\\midrule\multicolumn{7}{c}{\emph{Text Editing}}\\\midrule
  \% solved&\textbf{74\%} &43\% &30\%&33\%&0\%&4\%\\
  Solve time&  29s&65s&38s&80s&--&235s
